# Results

## Collaborative GIS Environment

A summary of background information layers loaded to the `background_layers.gpkg` geopackage of the `restoration_wedzin_kwa` project at the time of writing are included in Table \@ref(tab:tab-rfp-tracking).

```{r rfp-tracking-copy, eval = FALSE}
# first we will copy the doc from the Q project to this repo
q_path_stub <- "/Users/airvine/Projects/gis/restoration_wedzin_kwa/"

rfp_tracking_raw <- sf::st_read(paste0(q_path_stub, "background_layers.gpkg"),
            layer = "rfp_tracking",
            quite = TRUE)

# grab the metadata
md <- rfp_meta_bcd_xref()

# remove the `_vw` from the end of content
rfp_tracking_prep <- dplyr::left_join(
  rfp_tracking_raw %>% 
    dplyr::arrange(desc(timestamp)) %>% 
    dplyr::distinct(content, .keep_all = FALSE),
  
  md %>% 
    dplyr::select(content = object_name, url = url_browser, description),
  
  by = "content"
) %>% 
  arrange(content)

rfp_tracking_prep %>% 
  readr::write_csv("data/rfp_tracking_prep.csv")

```

```{r tab-rfp-tracking}
rfp_tracking_prep <- readr::read_csv(
  "data/rfp_tracking_prep.csv"
)

rfp_tracking_prep %>% 
  fpr::fpr_kable(caption_text = "Layers loaded to collaborative GIS project.",
                 footnote_text = "Metadata information for bcfishpass and bcfishobs layers can be provided here in the future but  currently can usually be sourced from https://smnorris.github.io/bcfishpass/06_data_dictionary.html .")



```

## Past Restoration Initiatives

Information amalgamated from past restoration initiatives has been added to the `restoration_wedzin_kwa` project in spatial formats through the `sites_restoration.gpkg` with details of source document locations, geometry type, description of the source document amalgamated, etc. documented in the `sites_restoration_gpkg_tracking.csv` file located [here](https://github.com/NewGraphEnvironment/restoration_wedzin_kwa_2024/blob/main/data/sites_restoration_gpkg_tracking.csv).  

<br>

### Works in Progress
@ncfdc1998 utilized detailed fish and fish habitat assessments to inform restoration planning in the Neexdzii Kwah
with numerous detailed prescriptions included within their reporting and digital deliverables. Unfortunately, the
digital deliverables could not be recovered.  Their work persists in pdf format however with summarized details of 
their prioritization work and prescription locations extracted from the pdf and included in the `ncfdc_1998_prescriptions_raw.csv` file located and viewable [here](https://github.com/NewGraphEnvironment/restoration_wedzin_kwa_2024/blob/main/data/ncfdc_1998_prescriptions_raw.csv).  The prioritization table for sub-basins in the region are included in Table \@ref(tab:tab-ncfdc_1998_priority). Next steps for this dataset include mining of prescription locations and addition to the csv so that it can be converted to a spatial file and added to the `restoration_wedzin_kwa` project through the `sites_restoration.gpkg`.

```{r tab-ncfdc_1998_priority}
ncfdc_1998_71a %>% 
  fpr::fpr_kable(caption_text = "Table 71a from the NCFDC 1998 report - summary of the prioritization of sub-basins in the Neexdzii Kwah region.")
```



